Equalizer\\Programming and User Guide

### 2012-04-10 Developer/ProgrammingGuide/paper.tex
Minor PG revision

\date{
  \vspace{14cm}
  \begin{textblock}{}(0,4.5)
    \hspace{-1cm}\includegraphics[width=22cm]{images/teaser.pdf}
  \end{textblock}
      {\Large The official reference for developing and
        deploying parallel, scalable OpenGL\texttrademark\ applications using
        the Equalizer parallel rendering framework}\\\vspace{1cm}
      Version 1.10 for Equalizer 1.2\\[\medskipamount]
  \today
}

@@

\clearpage
\pagenumbering{roman}
\pagestyle{headings}
\tableofcontents
%\clearpage
\listoffigures \vfill
\section*{Revision History}
{\center\begin{tabularx}{\textwidth}{|l|l|X|}
    \hline
    \bf Rev & \bf Date     & \bf Changes \\
    \hline
    1.0     & Oct 28, 2007 & Initial Version for Equalizer 0.4\\
    1.2     & Apr 15, 2008 & Revision for Equalizer 0.5\\
    1.4     & Nov 25, 2008 & Revision for Equalizer 0.6\\
    1.6     & Aug 07, 2009 & Revision for Equalizer 0.9\\
    1.8     & Mar 21, 2011 & Revision for Equalizer 1.0\\
    1.10    & Feb 17, 2012 & Revision for Equalizer 1.2\\
    \hline
  \end{tabularx}}
\clearpage

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/2D.pdf}
  {\caption{A 2D Compound}}
\end{wrapfigure}
2D decomposes the rendering in screen-space, that is, each contributing
rendering unit processes a tile of the final view. The recomposition
simply assembles the tiles side-by-side on the destination view. This
mode is also known as sort-first or SFR.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/DB.pdf}
  {\caption{\label{fDB}\small A Database Compound}}
\end{wrapfigure}
Volume rendering applications use an ordered alpha-based blending to
composite the result image. The depth buffer information is used to
composite the individual images correctly for polygonal data.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/EYE.pdf}
  {\caption{\label{fStereoCmp}\small A Stereo Compound}}
\end{wrapfigure}
Due to the frame consistency between the eye views, this modes scales very
well. The IO requirements for pixel transfer are small and constant. The number
of rendering resources used by stereo compounds is limited by the number of eye
passes, typically two.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \vspace{-2ex}
  \includegraphics[width=.618\textwidth]{images/DPlex.pdf}
  {\caption{ A DPlex Compound}}
\end{wrapfigure}
DPlex compounds assign full, alternating frames to individual rendering
units. The resulting images are copied to the destination channel, and
Equalizer load-balancing is used to ensure a steady framerate on the
destination window. This mode is also known as time-multiplex or AFR.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/tile.pdf}
  {\caption{\label{fTile}\small A Tile Compound}}
\end{wrapfigure}
Tile compounds have a low, constant IO overhead for the image transfers and can
provide good scalability when used with fill-limited applications. Tile
decomposition works transparently for all Equalizer-based applications.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \vspace{-2ex}\includegraphics[width=.618\textwidth]{images/Pixel.pdf}
  {\caption{\label{fPixel}\small A Pixel Compound}}\vspace{-8ex}
\end{wrapfigure}
Pixel compounds are similar to 2D compounds. While 2D compounds decompose the
screen space into contiguous regions, pixel compounds assign one pixel of a
regular kernel to each resource. The frusta of the source rendering units are
modified so that each unit renders an evenly distributed subset of pixels, as
shown in \fig{fPixel}\footnote{3D model courtesy of AVS, USA.}

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/Subpixel.pdf}
  {\caption{\label{fSubpixel}\small A Subpixel Compound}}
\end{wrapfigure}
Subpixel compounds are configured using the \textsf{subpixel} parameter, using
the values \textsf{[~index size ]}. The \textsf{index} parameter corresponds to
the current resource index and the \textsf{size} is the total number of
resources used.

@@

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/layout.png}
  {\caption{\label{fLayout}A Layout with four Views}}
\end{wrapfigure}
Enable the layout on each desired canvas by adding it to the
canvas. Since canvases reference layouts by name or index, layouts have
to be configured before their respective canvases in the configuration
file.

@@

\begin{figure}[ht!]\center
  \includegraphics[width=.9\textwidth]{images/wallLayout.jpg}
  {\caption{A Display Wall using a six-Segment Canvas with a two-View Layout}}
\end{figure}

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/monitorEq.pdf}
  {\caption{\label{fMonitorEq}Monitoring a Projection Wall}}
\end{wrapfigure}
A monitor equalizer adjusts the frame zoom of the output frames used to
observe the rendering, depending on the destination channel size. The
output frames are downscaled on the GPU before readback, which results
in optimal performance.

@@

\subsection{The Server}
The server may be started as a separate process or within the application
process. If it is started separately, simply provide the desired configuration
file as a parameter. It will listen on all network addresses, unless other
connection parameters are specified in the configuration file. If the
server is started within the application process, using the \textsf{--eq-config}
parameter, you will have to specify a connection description for the server in
the configuration file. Application-local servers do not create a listening
socket by default for security reasons.

@@

\subsection{The main Function}

@@

\subsection{The Application}

@@

\subsection{The Renderer}

@@

\section{The Programming Interface}

@@

\begin{description}
\item[eq] The core namespace for the Equalizer client library. The
  classes and their relationship in this namespace closely model the
  configuration file format. The classes in the \textsf{eq} namespace
  are the main subject of this Programming Guide.
\item[eq::util] The eq::util namespace provides common utility classes, which
  often simplify the usage of OpenGL functions. Most classes in this namespace
  are used by the Equalizer client library, but are usable independently
  from Equalizer for application development.
\item[eq::admin] The eq::admin namespace implements an administrative API to
  change the configurations of a running server. This admin API is not yet
  finalized and will very likely change in the future.
\item[eq::fabric] The eq::fabric namespace is the common data management and
  transport layer between the client library, the server and the administrative
  API. Most Equalizer client classes inherit from base classes in this namespace
  and have all their data access methods in these base classes.
\item[co] Collage is the network library used by Equalizer. It provides basic
  functionality for network communication, such as \textsf{Connection} and
  \textsf{ConnectionSet}, as well as higher-level functionality such as
  \textsf{Node}, \textsf{LocalNode}, \textsf{Object} and
  \textsf{Serializable}. Please refer to \sref{sNetwork} for an introduction
  into the network layer, and to \sref{sNetObject} for distributed objects.
\item[co::base] The co::base namespace provides C++ classes to abstract
  the underlying operating system and implements common helper
  functionality. One example is the \textsf{co::base::Clock} class
  providing a high-resolution timer. Classes in this namespace are fully
  documented in the API documentation on the Equalizer website, and are
  not subject of this Programming Guide.
\end{description}

Examples shipped with Equalizer are implemented in their own namespaces, e.g.,
\textsf{eqPly} or \textsf{eVolve}. They rely mostly on subclassing from the
\textsf{eq} namespace, with the occasional usage of functionality from the
\textsf{eq::util}, \textsf{eq::fabric}, \textsf{co} and \textsf{co::base}
namespace.

@@

{\footnotesize\lstinputlisting[firstline=89, lastline=98]
  {../../../equalizergraphics.com/documents/design/fileFormat.shtml}}

@@

{\footnotesize\lstinputlisting[firstline=100, lastline=109]
  {../../../equalizergraphics.com/documents/design/fileFormat.shtml}}

@@

{\footnotesize\lstinputlisting[firstline=111, lastline=129]
  {../../../equalizergraphics.com/documents/design/fileFormat.shtml}}

@@

{\footnotesize\lstinputlisting[firstline=130, lastline=139]
  {../../../equalizergraphics.com/documents/design/fileFormat.shtml}}

@@

{\footnotesize\lstinputlisting[firstline=141, lastline=162]
  {../../../equalizergraphics.com/documents/design/fileFormat.shtml}}

@@

{\footnotesize\lstinputlisting[firstline=163, lastline=172]
  {../../../equalizergraphics.com/documents/design/fileFormat.shtml}}

@@

{\footnotesize\lstinputlisting[firstline=176, lastline=182]
  {../../../equalizergraphics.com/documents/design/fileFormat.shtml}}

@@

{\footnotesize\lstinputlisting[firstline=184, lastline=187]
  {../../../equalizergraphics.com/documents/design/fileFormat.shtml}}

@@

{\footnotesize\lstinputlisting[firstline=188, lastline=209]
  {../../../equalizergraphics.com/documents/design/fileFormat.shtml}}

@@

{\footnotesize\lstinputlisting[firstline=211, lastline=236]
  {../../../equalizergraphics.com/documents/design/fileFormat.shtml}}

@@

{\footnotesize\lstinputlisting[firstline=238, lastline=259]
  {../../../equalizergraphics.com/documents/design/fileFormat.shtml}}

@@

{\footnotesize\lstinputlisting[firstline=262, lastline=333]
  {../../../equalizergraphics.com/documents/design/fileFormat.shtml}}

>>>>

\date{
  \vspace{14cm}
  \begin{textblock}{}(0,4.5)
    \hspace{-1cm}\includegraphics[width=22cm]{images/teaser.pdf}
  \end{textblock}
      {\Large The official reference for developing and
        deploying parallel, scalable OpenGL\texttrademark\ applications using
        the Equalizer parallel rendering framework}\\\vspace{1cm}
      Version 1.11.1 for Equalizer 1.2+git\\[\medskipamount]
  \today
}

@@

\clearpage
\pagenumbering{roman}
\pagestyle{headings}
\tableofcontents
%\clearpage
\listoffigures \vfill
\section*{Revision History}
{\center\begin{tabularx}{\textwidth}{|l|l|X|}
    \hline
    \bf Rev & \bf Date     & \bf Changes \\
    \hline
    1.0     & Oct 28, 2007 & Initial Version for Equalizer 0.4\\
    1.2     & Apr 15, 2008 & Revision for Equalizer 0.5\\
    1.4     & Nov 25, 2008 & Revision for Equalizer 0.6\\
    1.6     & Aug 07, 2009 & Revision for Equalizer 0.9\\
    1.8     & Mar 21, 2011 & Revision for Equalizer 1.0\\
    1.10    & Feb 17, 2012 & Revision for Equalizer 1.2\\
    1.11.1  & Apr 10, 2012 & Minor revision\\
    \hline
  \end{tabularx}}
\clearpage

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/2D.pdf}
  {\caption{2D Compound}}
\end{wrapfigure}
2D decomposes the rendering in screen-space, that is, each contributing
rendering unit processes a tile of the final view. The recomposition
simply assembles the tiles side-by-side on the destination view. This
mode is also known as sort-first or SFR.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/DB.pdf}
  {\caption{\label{fDB}Database Compound}}
\end{wrapfigure}
Volume rendering applications use an ordered alpha-based blending to
composite the result image. The depth buffer information is used to
composite the individual images correctly for polygonal data.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/EYE.pdf}
  {\caption{\label{fStereoCmp}Stereo Compound}}
\end{wrapfigure}
Due to the frame consistency between the eye views, this modes scales very
well. The IO requirements for pixel transfer are small and constant. The number
of rendering resources used by stereo compounds is limited by the number of eye
passes, typically two.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \vspace{-2ex}
  \includegraphics[width=.618\textwidth]{images/DPlex.pdf}
  {\caption{A DPlex Compound}}
\end{wrapfigure}
DPlex compounds assign full, alternating frames to individual rendering
units. The resulting images are copied to the destination channel, and
Equalizer load-balancing is used to ensure a steady framerate on the
destination window. This mode is also known as time-multiplex or AFR.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/tile.pdf}
  {\caption{\label{fTile}Tile Compound}}
\end{wrapfigure}
Tile compounds have a low, constant IO overhead for the image transfers and can
provide good scalability when used with fill-limited applications. Tile
decomposition works transparently for all Equalizer-based applications.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \vspace{-2ex}\includegraphics[width=.618\textwidth]{images/Pixel.pdf}
  {\caption{\label{fPixel}Pixel Compound}}\vspace{-8ex}
\end{wrapfigure}
Pixel compounds are similar to 2D compounds. While 2D compounds decompose the
screen space into contiguous regions, pixel compounds assign one pixel of a
regular kernel to each resource. The frusta of the source rendering units are
modified so that each unit renders an evenly distributed subset of pixels, as
shown in \fig{fPixel}\footnote{3D model courtesy of AVS, USA.}

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/Subpixel.pdf}
  {\caption{\label{fSubpixel}Subpixel Compound}}
\end{wrapfigure}
Subpixel compounds are configured using the \textsf{subpixel} parameter, using
the values \textsf{[~index size ]}. The \textsf{index} parameter corresponds to
the current resource index and the \textsf{size} is the total number of
resources used.

@@

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/layout.png}
  {\caption{\label{fLayout}Layout with four Views}}
\end{wrapfigure}
Enable the layout on each desired canvas by adding it to the
canvas. Since canvases reference layouts by name or index, layouts have
to be configured before their respective canvases in the configuration
file.

@@

\begin{figure}[ht!]\center
  \includegraphics[width=.9\textwidth]{images/wallLayout.jpg}
  {\caption{Display Wall using a six-Segment Canvas with a two-View Layout}}
\end{figure}

@@

\begin{wrapfigure}{l}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/monitorEq.pdf}
  {\caption{\label{fMonitorEq}Monitoring a Projection Wall}}
\end{wrapfigure}
A monitor equalizer adjusts the frame zoom of the output frames used to
observe the rendering, depending on the destination channel size. The
output frames are downscaled on the GPU before readback, which results
in optimal performance.

@@

\subsection{Server}
The server may be started as a separate process or within the application
process. If it is started separately, simply provide the desired configuration
file as a parameter. It will listen on all network addresses, unless other
connection parameters are specified in the configuration file. If the
server is started within the application process, using the \textsf{--eq-config}
parameter, you will have to specify a connection description for the server in
the configuration file. Application-local servers do not create a listening
socket by default for security reasons.

@@

\subsection{main Function}

@@

@@

\subsection{Renderer}

@@

\section{Programming Interface}

@@

\begin{description}
\item[eq] The core namespace for the Equalizer client library. The
  classes and their relationship in this namespace closely model the
  configuration file format. The classes in the \textsf{eq} namespace
  are the main subject of this Programming Guide.
\item[eq::util] The eq::util namespace provides common utility classes, which
  often simplify the usage of OpenGL functions. Most classes in this namespace
  are used by the Equalizer client library, but are usable independently
  from Equalizer for application development.
\item[eq::admin] The eq::admin namespace implements an administrative API to
  change the configurations of a running server. This admin API is not yet
  finalized and will very likely change in the future.
\item[eq::fabric] The eq::fabric namespace is the common data management and
  transport layer between the client library, the server and the administrative
  API. Most Equalizer client classes inherit from base classes in this namespace
  and have all their data access methods in these base classes.
\item[co] Collage is the network library used by Equalizer. It provides basic
  functionality for network communication, such as \textsf{Connection} and
  \textsf{ConnectionSet}, as well as higher-level functionality such as
  \textsf{Node}, \textsf{LocalNode}, \textsf{Object} and
  \textsf{Serializable}. Please refer to \sref{sNetwork} for an introduction
  into the network layer, and to \sref{sNetObject} for distributed objects.
\item[lunchbox] The lunchbox library provides C++ classes to abstract the
  underlying operating system and to implement common helper functionality for
  multi-threaded applications. Examples are \textsf{lunchbox::Clock} providing a
  high-resolution timer, or \textsf{lunchbox::MTQueue} providing a thread-safe,
  blocking fifo. Classes in this namespace are fully documented in the API
  documentation on the Equalizer website, and are not subject of this
  Programming Guide.
\end{description}

Examples shipped with Equalizer are implemented in their own namespaces, e.g.,
\textsf{eqPly} or \textsf{eVolve}. They rely mostly on subclassing from the
\textsf{eq} namespace, with the occasional usage of functionality from the
\textsf{eq::util}, \textsf{eq::fabric}, \textsf{co} and \textsf{lunchbox}
namespace.

@@

{\footnotesize\lstinputlisting[firstline=89, lastline=98]
  {../../../equalizergraphics/documents/design/fileFormat.shtml}}

@@

{\footnotesize\lstinputlisting[firstline=100, lastline=109]
  {../../../equalizergraphics/documents/design/fileFormat.shtml}}

@@

{\footnotesize\lstinputlisting[firstline=111, lastline=129]
  {../../../equalizergraphics/documents/design/fileFormat.shtml}}

@@

{\footnotesize\lstinputlisting[firstline=130, lastline=139]
  {../../../equalizergraphics/documents/design/fileFormat.shtml}}

@@

{\footnotesize\lstinputlisting[firstline=141, lastline=162]
  {../../../equalizergraphics/documents/design/fileFormat.shtml}}

@@

{\footnotesize\lstinputlisting[firstline=163, lastline=172]
  {../../../equalizergraphics/documents/design/fileFormat.shtml}}

@@

{\footnotesize\lstinputlisting[firstline=176, lastline=182]
  {../../../equalizergraphics/documents/design/fileFormat.shtml}}

@@

{\footnotesize\lstinputlisting[firstline=184, lastline=187]
  {../../../equalizergraphics/documents/design/fileFormat.shtml}}

@@

{\footnotesize\lstinputlisting[firstline=188, lastline=209]
  {../../../equalizergraphics/documents/design/fileFormat.shtml}}

@@

{\footnotesize\lstinputlisting[firstline=211, lastline=236]
  {../../../equalizergraphics/documents/design/fileFormat.shtml}}

@@

{\footnotesize\lstinputlisting[firstline=238, lastline=259]
  {../../../equalizergraphics/documents/design/fileFormat.shtml}}

@@

{\footnotesize\lstinputlisting[firstline=262, lastline=333]
  {../../../equalizergraphics/documents/design/fileFormat.shtml}}
